Applying Deep Learning to Acoustic Analysis in the PICO-60 Dark Matter Detector

### 2018-08-11 documentation/paper.tex
Cleared up description of semi-supervised learning

\begin{enumerate}
    \item Determine whether machine learning techniques can function comparably to a conventionally derived discriminator. Furthermore, develop several neural network architectures and apply them to different input formats, in order to find which are the most effective.

    \item Determine whether these techniques can be extended to develop an effective discriminator function based on incomplete and inaccurate initial information.

@@

@@

During early tests with variations on $DeepConv(\omega)$, it was revealed to have an incredible capacity to separate run types based on unwanted biases. It was able to separate the AmBe calibration run from combined background runs with 96\% accuracy, and to separate all calibration runs from background runs with 91\% accuracy. This was determined to be caused by the network detecting and fitting on biases in the white noise at the beginning and hydraulic recompression sounds at the end of the raw waveform. This extraneous information was subsequently cropped out.

@@

Throughout a variety of different network architectures, the best validation accuracy obtained was a deceptively high 85\%. The corresponding training accuracy of 100\% indicates that severe overfitting is taking place. Also, the class-wise standard deviation was a very high 0.95, which implies its decisions are nearly random (as is visually apparent in \ref{image_window}). It is likely fitting on some form of noise within $\iota$. The fact that throughout many trials (including a grid search), no good performance on validation data was observed, provides significant evidence that images provide insufficient information for effective discrimination.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{image_window}
    \caption{\label{image_window} Accuracy of $ImageConv(\iota)$ compared to Acoustic Parameter at discriminating between run types (85\% accuracy).}
\end{figure}

@@

Semi-supervised learning is the idea of training a machine learning model on a set of labeled data in addition to a set of unlabeled data. This has the advantage of requiring fewer labels to be collected than conventional supervised learning. However, it also has a very powerful advantage when training data is impure and the function to be learned is relatively simple (not simple in that it can be easily calculated by a person, but simple in that it does not require a huge set of training data). It allows a neural network to reinforce its own decisions.

@@

>>>>

\begin{enumerate}
    \item Determine the most effective machine learning architecture to use in development of a discriminator. Furthermore, develop several neural network architectures and apply them to different input formats, in order to find which are the most effective.

    \item Determine whether these techniques can be extended, applying novel algorithms based on semi-supervised learning, to develop an effective discriminator function based on incomplete and inaccurate initial information.

@@

For all supervised learning configurations, accuracy and class-wise standard deviation statistics were computed using the calibration run type as a base line, as opposed to Acoustic Parameter.

Because supervised learning systems are trained strictly to replicate ground truth data provided to them (the run type), their predictions will tend to incorporate biases similar to those found in the training set. Thus, comparison to Acoustic Parameter is not particularly meaningful.

@@

During early tests with variations on $DeepConv(\omega)$, it was revealed to have a surprising and undesirable capacity to separate run types based on meaningless information. It was able to separate the AmBe calibration run from combined background runs with 96\% accuracy, and to separate all calibration runs from background runs with 91\% accuracy. This was determined to be caused by the network detecting and fitting on biases in the white noise at the beginning and hydraulic recompression sounds at the end of the raw waveform. This extraneous information was subsequently cropped out.

@@

Throughout a variety of different network architectures, the best validation accuracy obtained was a deceptively high 85\%. The corresponding training accuracy of 100\% indicates that severe overfitting is taking place. Also, the class-wise standard deviation was a very high 0.95, which implies its decisions are nearly random. It is likely fitting on some form of noise within $\iota$. The fact that throughout many trials (including a grid search), no good performance on validation data was observed, provides significant evidence that images provide insufficient information for effective discrimination.

@@

@@

Semi-supervised learning is an uncommon machine learning technique, relative to widely-used supervised and unsupervised learning. The concept is to train a machine learning model on a set of labeled data in addition to a set of unlabeled data.

In the context of PICO-60, the labeled sets consist of single-bubble events from the calibration sets, which are known to contain approximately 90\% neutrons, and the background radiation data sets, which contain approximately 99\% alpha particles. In general, these labeled sets should have a strong but not perfect correlation with particle types. The unlabeled data can consist of any mixture of particle types, as long as they are all represented in the labeled data.

The primary advantage of this technique is clear: it requires fewer labels to be collected than conventional supervised learning. However, it also has a very powerful advantage when training data is impure and the function to be learned is relatively simple (not simple in that it can be easily calculated by a person, but simple in that it does not require a huge set of training data): It allows a neural network to reinforce its own decisions.

@@

Since a goal of these techniques is for the network to reinforce the pattern of particle types which is represented imperfectly in the run types, it is reasonable to compare them to Acoustic Parameter. Thus, the accuracy and class-wise standard deviation use Acoustic Parameter as a baseline.
